---
title: "Reading group"
---
```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE,
  fig.align = "center",
  cache = F
)

```
<style>
  table th, table td {
    white-space: nowrap;
  }
  table th:nth-child(1), table td:nth-child(1) { width: 2cm; }
  table th:nth-child(2), table td:nth-child(2) { width: 5cm; }
  table th:nth-child(3), table td:nth-child(3) { width: 2cm; }
  table th:nth-child(4), table td:nth-child(4) { width: 2cm; }
  table th:nth-child(5), table td:nth-child(5) { width: 2cm; }
</style>

Each GLE$^2$N meeting begins with a discussion of a recent piece of relevant literature. A lead discussant typically prepares slides summarising the article, which serve as the basis for the group discussion. To facilitate ongoing dialogue, the lead discussant also compiles a mock review document highlighting the key discussion points. The article, slides, and mock review document are provided below. All discussions take place online via Zoom.

## Reading list 📝
```{r, echo = F, eval=T}
time = c('21/03/25<br>13:00-14:00',
         '14/02/25<br>13:00-14:00',
         '15/11/24<br>13:00-14:00<br>',
         '18/10/24<br>13:00-14:00<br>')
article = c(
  'Gnecco, N., Merga, E. and Engelke, S. (2024). <a href="https://drive.google.com/file/d/1M2WjpOO0moccyxrLhs-Nq3y9_lkP8rgK/view?usp=sharing" target="_blank">Extremal Random Forests</a>. <em>Journal of the <br>American Statistical Association</em>, 119(548), pp.3059–3072, 82–95.',
  'Huser, R. & Wadsworth, J. (2019). <a href="https://drive.google.com/file/d/1wj2v60i9AZSbmcMRIJ1C59dCxnHyTxlP/view?usp=sharing" target="_blank">Modeling Spatial Processes with Unknown Extremal<br>Dependence Class</a>, <em>Journal of the American Statistical Association</em>, 114(525), 434–444.',
  'Li, R., Leng, C., & You, J. (2020). <a href="https://drive.google.com/file/d/1pnrSgfoF4xaHRdhWDoWHGL7OfRWrgyos/view?usp=sharing" target="_blank">Semiparametric Tail Index Regression</a>. <em>Journal of Business</em><br>&<em>Economic Statistics</em>, 40(1), 82–95.',
  
  'Olafsdottir, H. K., Rootzén, H., & Bolin, D. (2021). <a href="https://drive.google.com/file/d/1SDMxewRKqTujifQkXDTLoerxDuoVDrPI/view?usp=sharing" target="_blank">Extreme rainfall events in the<br>Northeastern United States become more frequent with rising temperatures, but their<br>intensity distribution remains stable</a>. <em>Journal of Climate</em>, 34(22), 8863-8877.')

discussant = c('Lambert de Monte',
               'Chenglei Hu',
               'Johnny Lee',
               'Daniela Castro-Camilo')

slides = c('slides4',
           '<a href="https://drive.google.com/file/d/1qfLUutAMtiIglgiiLFpF1vdIhyiXdf2_/view?usp=drive_link" target="_blank">slides3</a>',
           
           '<a href="https://drive.google.com/file/d/1ziUE0SIHDDKG7yPHvJHgDI385KFOZMrW/view?usp=sharing" target="_blank">slides2</a>',
           
           '<a href="https://drive.google.com/file/d/1ilwjBS_KVJZxWJaiSYAFYU4_jJovFDEA/view?usp=sharing" target="_blank">slides1</a>')

mocks = c('mockrev4',
          '<a href="https://docs.google.com/document/d/1wEb9aU256I0UwVzR1TztJB2y0NINnlmBc11UtPP9pO8/edit?usp=sharing" target="_blank">mockrev3</a>',
          
          '<a href="https://docs.google.com/document/d/1W8JH74O6Slj3ciDQSUdS5nKQ87ZjdfEzcvzHFsSzU4o/edit?usp=sharing" target="_blank">mockrev2</a>',
          
          '<a href="https://docs.google.com/document/d/1biNc5KB6N8HhenZgOqCDYQBZvIHmNwtfTSOy65XojY0/edit?usp=sharing" target="_blank">mockrev1</a>')


d = data.frame(time, article, discussant, slides, mocks)


kbl(d, escape = F, col.names = c('Date and time', 'Article', 'Discussant', 'Slides', 'Mock reviews'), 
    align = c('c','l','c','c','c')) %>%
  kable_styling(htmltable_class = 'lightable-hover', full_width = T)


```


